---
title: "NYC severe Fire Incidents Analysis"
author: "Ujjawal Gupta"
date: "March 20, 2018"
output:
html_document: yes
keep_md: yes
---

``````{r Setup, include=FALSE, results='hide', warning=FALSE}
library(knitr)

opts_chunk$set(cache.path="cache/",
                   cache=FALSE,
                   echo=TRUE,
                   message=FALSE,
                   warning=FALSE)
```

```{r import libs, echo = FALSE}
library(tidyverse)
library(readr)
library(ggthemes)
library(sqldf)
library(leaflet)
library(jsonlite)
library(RColorBrewer)
library(sqldf)
library(geosphere)
library(sp)
library(rgeos)
library(plyr)
library(fossil)
```
## 1. Location of Severe Fires

The map used is Stamen.Toner, There were few irregularities in the data.

- 16 tuples did not had any Longitude and Latitude Values.

- 7 Longitude and Latitude Values of the Incidents were wrong, i.e Data set described the events in Manhattan but the Geo Location was pointed outside of NYC.

Such Values were filtered using using NYC boundries fro North, South, East and West Direstions West -74.257159 East -73.699215 North 40.915568 South 40.495992.

```{r Q1, echo=FALSE}
severe_incident <- read.csv("severe_incidents.csv", header = TRUE, stringsAsFactors=FALSE)

severe_inc <- sqldf("Select * from severe_incident where Longitude < '-73.699215' and Longitude > '-74.257159' and Latitude > '40.495992' and Latitude < '40.915568'")

popup1 <- paste("Incident Type:",severe_inc$INCIDENT_TYPE_DESC,"<br/>",
                "When:",severe_inc$INCIDENT_DATE_TIME,"<br/>",
                "Property Type:",severe_inc$PROPERTY_USE_DESC,"<br/>",
                "Action Taken:",severe_inc$ACTION_TAKEN1_DESC,"<br/>")

Ans1 <- leaflet(severe_inc) %>%
        addProviderTiles("Stamen.TonerLite",
                         options = providerTileOptions(attribution = ""))%>%
        setView(lng = -74.0156491, lat = 40.7022541, zoom = 10)%>%
       addCircleMarkers(lng = ~Longitude, lat = ~Latitude, col = "red", radius = 1, popup = popup1, stroke = 0.1, opacity = 0.7) %>%
  addEasyButton(easyButton(
    icon="fa-globe", 
    title="New York City",
    onClick=JS("function(btn, map){ map.setZoom(10); }")))

Ans1
```

## 2. Layers and Clusters
###  2a. Color by Type of Property

There were Many Categories related to Type of Property.They were collapsed into 2 Broader categories.


 - **Residential/Multihousing** The Category number Ranging from 400 to 499 was categories into this group. 
 
 - **Commercial/Others** Rest all were added in this category.

Residential/Multihousing property are those with sole residential purpose, any other property with any commercial/business use were categorised in Category 2.
```{r Q2, echo=FALSE}
pal = colorFactor("Set1", domain = severe_inc$PROPERTY_USE_DESC)
color1_Property = pal(severe_inc$PROPERTY_USE_DESC)


Ans2a <- Ans1%>%
  addCircleMarkers(lng = ~Longitude, lat = ~Latitude, col = color1_Property, radius = 1, popup = popup1, stroke = 0.1, opacity = 0.7)%>%
addLegend(pal = pal, values = ~severe_inc$PROPERTY_USE_DESC, title = "Property Types",position="bottomright")

Ans2a
```

### 2b. Cluster

Added marker clustering, so that zooming in will reveal the individual locations but the zoomed out map only shows the clusters.

```{r Q3, echo = FALSE}

Ans2b <- leaflet(severe_inc) %>%
  addProviderTiles("Stamen.TonerLite",
                   options = providerTileOptions(attribution = ""))%>%
  setView(lng = -74.0156491, lat = 40.7022541, zoom = 10)%>%
  addCircleMarkers(lng = ~Longitude, lat = ~Latitude, col = color1_Property, radius = 5, popup = popup1, stroke = 2, opacity = 0.7,clusterOptions = markerClusterOptions())%>%
  addLegend(pal = pal, values = ~severe_inc$PROPERTY_USE_DESC, title = "Property Types",position="bottomright") %>%
  addEasyButton(easyButton(
    icon="fa-globe", 
    title="New York City",
    onClick=JS("function(btn, map){ map.setZoom(10); }")))

Ans2b

```


## 3. Fire Houses

1. The Data for Fire Houses had 5 values with Long, Lat values not available, those values were removed. A Fire icon is used to denote the location Fire Stations on the Map.

2. Units on Scene Parameter is used to adjust size of circle Markers by severity.


```{r Q4, echo= FALSE}


  Firehouses1 <- read.csv("FDNY_Firehouse_Listing.csv", header = TRUE, stringsAsFactors=FALSE)

Firehouses <- sqldf("Select * from Firehouses1 where Longitude < '-73.699215' and Longitude > '-74.257159' and Latitude > '40.495992' and Latitude < '40.915568'")

FireIcons <- icons(
  iconUrl = "fire.png",
  iconWidth = 15, iconHeight = 15,
  iconAnchorX = 7.5, iconAnchorY = 7.5)


Ans3 <- leaflet() %>%
  addProviderTiles("Stamen.TonerLite",options = providerTileOptions(attribution = "")) %>%
  addCircleMarkers(data=severe_inc,lng = ~Longitude, lat = ~Latitude, radius=~UNITS_ONSCENE/4, fillOpacity=0.7,
  popup=~paste("Incident on:",severe_inc$STREET_HIGHWAY,"<br>",
               "Units on Scene:",severe_inc$UNITS_ONSCENE,"<br/>"),group = "Incidents") %>%
  addMarkers(data=Firehouses,lng = ~Longitude, lat = ~Latitude,
             icon = FireIcons,
             popup=~paste("Facility Address:",Firehouses$FacilityAddress,"<br>",
                          "Borough:",Firehouses$Borough,"<br/>"),group ="Firehouses")%>%
  setView(lng = -74.0156491, lat = 40.7022541, zoom = 10)%>%
  addLayersControl(overlayGroups = c("Incidents","Firehouses"),
    options = layersControlOptions(collapsed = TRUE))

Ans3
```

## 4. Distance from Firehouse and Response Time
###4a. Calculate Distance

The distance to the Nearest Firestation is Calculated in 2 steps.

1. Using an R function The BIN value of the Nearest Fire Station is calculated for each Incident. This is merged with Firehouse data to find corresponding Location of the Incident v Firestation.

2. Using distHaversine funtion the distance was calculated and then value was merged to Severe Incidents data

The plot shows distance in miles and response time in minutes, the size of the circles depends on the severity of the Incidents. I.e more units on scene, larger is the circle.


```{r Q5, echo= FALSE}
severe_inc$Incident_time <- as.POSIXct(severe_inc$INCIDENT_DATE_TIME,format="%m/%d/%Y %H:%M")

severe_inc$Arrival_time <- as.POSIXct(severe_inc$ARRIVAL_DATE_TIME,format="%m/%d/%Y %H:%M")

severe_inc$Duration <- difftime(severe_inc$Arrival_time,severe_inc$Incident_time,units = "mins")

A <- matrix(c(severe_inc$Longitude,severe_inc$Latitude), ncol=2)
B <- matrix(c(Firehouses$Longitude,Firehouses$Latitude), ncol=2)


C <- matrix(c(distHaversine(B,A, r=6378137)),nrow=218)

UniqueCoordinates <- data.frame(severe_inc[,26:25])
UniqueCoordinates$Id <- formatC((1:nrow(UniqueCoordinates)), width=3,flag=0)


## Generate a function that looks for the closest Firestation for each id coordinates
NearestFS <- function(id){
  tmp <- UniqueCoordinates[UniqueCoordinates$Id==id, 1:2]
  Firetemp <- rbind(tmp,Firehouses[,6:5])
  BIN <- earth.dist(Firetemp, dist=TRUE)[1:(nrow(Firetemp)-1)]
  BIN <- which.min(BIN)
  BIN <- Firehouses[BIN,10]
  BIN <- data.frame(BIN, Firetemp=tmp)
  return(BIN)
}

#apply to each id and the merge
CoordinatesFS <- ldply(UniqueCoordinates$Id, NearestFS)


Distancetemp <- sqldf("Select * from CoordinatesFS sev join Firehouses fire on sev.BIN = fire.BIN")

A <- matrix(c(Distancetemp$Longitude,Distancetemp$Latitude), ncol=2)
B <- matrix(c(Distancetemp$Firetemp.Longitude,Distancetemp$Firetemp.Latitude), ncol=2)

C <- matrix(c(distHaversine(A,B, r=6378137)))

severe_inc$DistanceNearestFS <- C

Ans4a <- ggplot(data = severe_inc, aes(x=DistanceNearestFS*0.00062137, y=Duration, color=DistanceNearestFS*0.00062137, alpha = 0.7 ))+
  geom_point(shape = 16, size = severe_inc$UNITS_ONSCENE/5, show.legend = FALSE) +
  geom_smooth(method="loess", se=F,show.legend = FALSE) + 
  xlim(c(0, 3)) + 
  ylim(c(0, 75)) +
  scale_color_gradient(low = "#0091ff", high = "#f0650e")+
  theme_bw() +
  xlab("Nearest Firestation Distance(in Miles)")+
  ylab("Response Time")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  ggtitle("Fire Station Distance vs Response time")

Ans4a
```


### 4b. Map of Response Times

Response time was calculated using Incindet time and Arrival time, this difference is used as the radius of individual circles. Further the color is based on the property type for the incidents.

```{r Q6, echo= FALSE}
Ans4b <- leaflet(severe_inc) %>%
  addProviderTiles("Esri.WorldStreetMap",options = providerTileOptions(attribution = "")) %>%
  addCircleMarkers(lng = ~Longitude, lat = ~Latitude, radius= ~Duration/4, fillOpacity=0.7,col = color1_Property, popup=~paste("Distance Nearest Firestation(m):",severe_inc$DistanceNearestFS,"<br>","Response time(mins):",severe_inc$Duration,"<br/>"))%>%
  addLegend(pal = pal, values = ~severe_inc$PROPERTY_USE_DESC, title = "Property Types",position="bottomright")%>%
setView(lng = -74.0156491, lat = 40.7022541, zoom = 10)%>%
  addEasyButton(easyButton(
    icon="fa-globe", 
    title="New York City",
    onClick=JS("function(btn, map){ map.setZoom(10); }")))

Ans4b

```

