---
title: "Winter Olympics"
author: "Ujjawal Gupta"
date: "February 20, 2018"
output:
  pdf_document: yes
  html_document: yes
keep_md: yes
---

``````{r Setup, include=FALSE, results='hide', warning=FALSE}
library(knitr)

opts_chunk$set(cache.path="cache/",
                   cache=FALSE,
                   echo=TRUE,
                   message=FALSE,
                   warning=FALSE)
```

```{r import libs, echo = FALSE}
library(tidyverse)
library(readr)
library(ggthemes)
library(sqldf)
library(manipulateWidget)
library(ggplot2)
library(gridExtra)
library(rvest)
library(stringr)
library(plotly)
library(DT)
```
## 1. Medal Counts over Time

For this task I have calculated and arranged the top countries based on (Gold, Silver and Bronze)medals won from 1924 to 2014.
- We find out that Canada has won most number of Gold medals, USA most number of Silver medals and Finland Most number of Bronze Medals.

- Different IOC names of Germany were merged together
- Russian Union Divided into 13 countries in 1992, so the IOC code URS for sovient Union has not been merged with Russia code RUS. Data for both of these codes have been represented individually.

```{r Q1, echo=FALSE}
Winter <- read.csv("winter.csv")

curr_country <- read.csv("dictionary.csv")

teams <- merge(x = Winter, y = curr_country, by = "Code", all = TRUE)


totalGold <- sqldf("select code, count(Medal)as totalG from teams where Medal = 'Gold' group by code")

totalSilver <- sqldf("select code, count(Medal)as totalS from teams where Medal = 'Silver' group by code")

totalBronze <- sqldf("select code, count(Medal)as totalB from teams where Medal = 'Bronze' group by code")

totalMedal <- merge(x = curr_country, y = totalGold, by = "Code", all = TRUE)

totalMedal <- merge(x = totalMedal, y = totalSilver, by = "Code", all = TRUE)

totalMedal <- merge(x = totalMedal, y = totalBronze, by = "Code", all = TRUE)

#Ques 1
 
  TempGold <- count(totalMedal,Code) %>% filter(totalMedal$totalG > 25)
  TGold <- filter(totalMedal, Code %in% TempGold$Code)
  
  subplot1 <- ggplot(data = TGold, aes(x = reorder(Code, -totalG), y = totalG, na.rm = FALSE))+
    geom_bar(stat= "identity",show.legend = FALSE, fill = "Gold")+
    xlab("")+
    ylab("Total number of Medals")+
  theme_bw()+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),axis.text.x = element_text(angle=90, size=rel(0.8), hjust=1))+ ggtitle("Top Gold Medal Winners")
  subplot1

  TempSilver <- count(totalMedal,Code) %>% filter(totalMedal$totalS > 25)
  TSilver <- filter(totalMedal, Code %in% TempSilver$Code)
  
  subplot2 <- ggplot(data = TSilver, aes(x = reorder(Code, -totalS), y = totalS, na.rm = FALSE))+
    geom_bar(stat= "identity",show.legend = FALSE, fill = "#bdbdbd")+
    xlab("")+
    ylab("")+
    theme_bw()+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),axis.text.x = element_text(angle=90, size=rel(0.8), hjust=1))+ ggtitle("Top Silver Medal Winners")
  subplot2
  
  TempBronze <- count(totalMedal,Code) %>% filter(totalMedal$totalB > 25)
  TBronze <- filter(totalMedal, Code %in% TempBronze$Code)
  
  subplot3 <- ggplot(data = TBronze, aes(x = reorder(Code, -totalB), y = totalB, na.rm = FALSE))+
    geom_bar(stat= "identity",show.legend = FALSE, fill = "#f03b20")+
    xlab("")+
    ylab("")+
    theme_bw()+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),axis.text.x = element_text(angle=90, size=rel(0.8), hjust=1))+ ggtitle("Top Bronze Medal Winners")
  subplot3
  
  Ans1 <- grid.arrange(subplot1, subplot2, subplot3, ncol=3, top = "Medal Distribution Winter olympics 1924 - 2014")
```

## 2. Medal Counts adjusted by Population, GDP

For this I chose Points = 3(Gold) + 2(Silver) + 1(Bronze) as valid method to Rank Teams.


 - We find out that **Canada** has most number of points but when we consider points won per GDP of the country **Russia** comes out on top. 
 
 - **Lietchtenstein** is the top country we consider Points per population due to its low 37,000 population.

```{r Q2, echo=FALSE}
totalMedal$Points <- ((totalMedal$totalG*3)+(totalMedal$totalS*2)+(totalMedal$totalB*1))
  
  totalMedal$GDPpPoints <- (totalMedal$Points/totalMedal$GDP.per.Capita)
  
  TempPoints <- count(totalMedal,Code) %>% filter(totalMedal$Points > 500)
  TPoints <- filter(totalMedal, Code %in% TempPoints$Code)
  
  plot2 <- ggplot(data = TPoints, aes(x = reorder(Code, -Points), y = Points, na.rm = FALSE))+
    geom_bar(stat= "identity",show.legend = FALSE, fill = "Blue" )+
    xlab("")+
    ylab("Points")+
    theme_bw()+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),axis.text.x = element_text(angle=90, size=rel(0.8), hjust=1))+ ggtitle("Most Successful Countries based on Points", subtitle = "Points = Gold*3 + Silver*2 + Bronze*1")
  plot2
  
  TempGDP <- count(totalMedal,Code) %>% filter(totalMedal$GDPpPoints > 0.01)
  TGDP <- filter(totalMedal, Code %in% TempGDP$Code)
  
  plot3 <- ggplot(data = TGDP, aes(x = reorder(Code, -GDPpPoints), y = GDPpPoints, na.rm = FALSE))+
    geom_bar(stat= "identity",show.legend = FALSE, fill = "Gold" )+
    xlab("")+
    ylab("Points/GDP per capita")+
    theme_bw()+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),axis.text.x = element_text(angle=90, size=rel(0.8), hjust=1))+ ggtitle("Most Successful Countries based on Points/GDP per capita")
  plot3
  
  totalMedal$PoppPoints <- (totalMedal$Points/totalMedal$Population)
  
  TempPop <- count(totalMedal,Code) %>% filter(totalMedal$PoppPoints > 0)
  TPop <- filter(totalMedal, Code %in% TempPop$Code)
  
  plot4 <- ggplot(data = TPop, aes(x = reorder(Code, -PoppPoints), y = PoppPoints, na.rm = FALSE))+
    geom_bar(stat= "identity",show.legend = FALSE, fill = "Green" )+
    xlab("")+
    ylab("Points/Population")+
    theme_bw()+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),axis.text.x = element_text(angle=90, size=rel(0.8), hjust=1))+ ggtitle("Most Successful Countries based on Points/Population")
  plot4
```

## 3. Host Country Advantage

If a country won more percentage of medals(Medals won by country in that year/Total number of medals in that year) in its host year as compared to other terms, then we can say that country was with home advantage.

But we see that on one hand countries can be said as they enjoy home advantage, it is not true for all the host nations.

 - **USA** Hosted Winter Olympics in 1932,1960,1980,2002 so it can be seen by there trend that they had more share of total medals everytime they hosted the olympics as compared to previous term. Most noticebly in 1932 when they won aroung 40% of the total medals.
 
 - **Germany** Hosted Olympics in 1936 but in that year they had only 6.4 percent of the total medal share, whereas in the year 1976 when Austria Hosted the Olympics they has 30.4 percent of the share.

```{r Q3, echo = FALSE}

#Ques 3
  
  wiki_hosts <- read_html("https://en.wikipedia.org/wiki/Winter_Olympic_Games")
  hosts <- html_table(html_nodes(wiki_hosts, "table")[[5]], fill=TRUE)
  hosts <- hosts[-1,1:3]
  hosts$city <- str_split_fixed(hosts$Host, n=2, ",")[,1]
  hosts$country <- str_split_fixed(hosts$Host, n=2, ",")[,2]
  
  HostCountry <- merge(x = Winter, y = hosts, by = "Year", all = TRUE)
  
  HostCountry$city <- NULL
  
  HostMedals <- sqldf("select Year,code,country, count(Medal)as totalH from HostCountry group by Year,code,country")
  
  HostMedals1 <- sqldf("select Year, Count(Medal)as Total from HostCountry group by Year")
  
  HostMedals <- merge(x = HostMedals, y = HostMedals1, by = "Year", all = TRUE)
  
  HostMedals$Percentage <- (HostMedals$totalH/HostMedals$Total)*100
  
  curr_country <- read.csv("dictionary.csv")
  
  curr_country$Hostcountry <- curr_country$Country
  
  HostMedals <- merge(x = HostMedals, y = curr_country, by = "Code", all = TRUE)
  
  HostMedals$Population <- NULL
  
  HostMedals$GDP.per.Capita <- NULL
  
  HostMedals$Country <- NULL
  
  TempHost <- read.csv("Host.csv")
  THostMedals <- filter(HostMedals, Code %in% TempHost$Code)
  
  plot5 <- ggplot(data=THostMedals, aes(colour = Hostcountry, x=Year, y=Percentage, group = Hostcountry)) + 
    geom_point(size = 1.5) +
    geom_line()+
    ylab("Percentage of Medals Won") +
    xlab("") + 
    theme_bw()+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),axis.text.x = element_text(angle=90, size=rel(0.8), hjust=1))+ ggtitle("Host Nations Medal Percentage share trend", sub = "France 1924,1968,1992 Switzerland 1928,1948 USA 1932,1960,1982,2002 Germany 1936 Norway 1952,1994 Italy 1956,2006 Austria 1964,1976 Japan 1972,1998 Canada 1988,2010 Russia 2014") 
  
  plot5

```


## 4. Country success by sport / discipline / event

1. Most number of medals won in Ice hockey is by Canada 14 Mens and 5 Womens Medals, closely followed by United States With 10 Mens and 5 Womens Medals.

2. Largest number of medals won in any Sport is by Norway in Skiing more than 200 medals.

**Bonus Point : In making this graph, data wrangling has been done as such that multiple medal distributed in team events are counted as one.**

```{r Q4, echo= FALSE}


  Event <- sqldf("Select Year,Sport,Event,Code,Medal,Gender,Count(Medal)as MedalsWon from Winter group by Year,Sport,Event,Code,Medal,Gender")
  
  Event$MedalsWon <- NULL
  
  Eventdis <- sqldf("Select Code,Sport, Gender, Count(Medal) as Countwon
                     from Event group by Code,Sport,Gender")
  
  Icehockey <- sqldf("Select Code, Gender,Countwon
                     from Eventdis where Sport = 'Ice Hockey'")
  
  Ans4 <- ggplot(Icehockey, aes(fill=Gender, y=Countwon, x= reorder(Code, -Countwon))) +
    geom_bar(stat="identity")+xlab("")+
    ylab("Total Medals Won")+
    theme_bw()+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),axis.text.x = element_text(angle=90, size=rel(1), hjust=0.5))+ ggtitle("Most Successful countries in Ice Hockey Winter Olympics")
  
  Ans4
  
  Eventtotal <- sqldf("Select Code,Sport, Count(Medal) as Countwon
                     from Event group by Code,Sport")
  
  Totalmed <- sqldf("Select Code, Count(Medal) as Countwon
                     from Event group by Code order by Countwon DESC limit 10")
  Top10 <- filter(Eventtotal, Code %in% Totalmed$Code)
  
  Ans4_2 <- ggplot(Top10, aes(fill=Sport, y=Countwon, x= reorder(Code, -Countwon))) +
    geom_bar(stat="identity")+xlab("")+
    ylab("Total Medals Won")+
    theme_bw()+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),axis.text.x = element_text(angle=90, size=rel(1), hjust=0.5))+ ggtitle("Sportwise Medal Distribution of 10 Most Successful countries")
  
  Ans4_2

```

## 5. Most successful athletes

1. Most Succesfull male athlete of winter olympics is Ole Einar from Norway.

2. Most Successful female athlete of all time in winter olympics is Stefania Belmondo who participated in skiing.


```{r Q5, echo= FALSE}
#Ques 5
  SuccessAthlete <- sqldf("Select Athlete,Gender,Sport,Code,Count(Athlete)as MedalsWon from Winter group by Athlete")
  
   SuccessMale <- sqldf("Select Athlete,Gender,Sport,Code,MedalsWon from SuccessAthlete where Gender='Men' order by MedalsWon DESC limit 10")
   
   SuccessFemale <- sqldf("Select Athlete,Gender,Sport,Code,MedalsWon from SuccessAthlete where Gender='Women' order by MedalsWon DESC limit 10")
   
   plot6 <- ggplot(data = SuccessMale, aes(x = reorder(Athlete, -MedalsWon), y = MedalsWon)) +
     geom_segment( aes(x= reorder(Athlete, -MedalsWon), xend=Athlete, y=0, yend=MedalsWon, color = Code, size = 0.1)) +
     geom_point( size=4, color="Black", fill=alpha("Black", 0.3), alpha=0.5, shape=21, stroke=1, show.legend = FALSE)+
     xlab("")+
     ylab("Total Medals Won")+
     theme_bw()+
     theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),axis.text.x = element_text(angle=90, size=rel(1), hjust=0.5))+ ggtitle("Top 10 Most Successful Male Athletes of Winter Olympics")
   
   plot6
   
   plot7 <- ggplot(data = SuccessFemale, aes(x = reorder(Athlete, -MedalsWon), y = MedalsWon, fill = Sport)) +
     geom_segment( aes(x= reorder(Athlete, -MedalsWon), xend=Athlete, y=0, yend=MedalsWon, color = Sport, size = 0.1)) +
     geom_point( size=4, color="Black", fill=alpha("Black", 0.3), alpha=0.5, shape=21, stroke=1)+
     xlab("")+
     ylab("Total Medals Won")+
     theme_bw()+
     theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),axis.text.x = element_text(angle=90, size=rel(1), hjust=0.5))+ ggtitle("Top 10 Most Successful Female Athletes of Winter Olympics")
   
   plot7

```


## 6. Making plots interactive

I choose to make following plots interactive:

 - **Host Nations Medal Percentage share trend** : The Graph itself has line trend for all the host nation till 2014. So making it interactive will enable user to see trend of a particular host nation with much less distraction.
 - **Sportwise Medal Distribution of Countries** : The plot showed cumulative distribution of medal count of countries. By making it interactive user can find additional information like medal count of those countries in individual Event.

```{r Q6, echo= FALSE}
#Ques 6
   
  library(plotly)
  
  ggplotly(plot5)
  
  ggplotly(Ans4_2)
  

```

## 7. Data Table

This data Table can be used to find out things like which country won a medal in a particlar event in winter olympics or How many events USA has medal in etc.


```{r Q7, echo= FALSE}

table1 <- Event %>%
    select(Year, Sport, Event, Gender, Code, Medal) %>%
    arrange(Year,Sport)  %>%
    datatable(rownames = FALSE,
              colnames = c("Year", "Sport", "Subcategory", "Participant", "Winning_Country", "Medal"), 
              caption =  "Winter Olympics Winners 1924 - 2014", 
              filter = "top")
  
  table1

```